---
title: "An overview of oolax"
author: "Paul Northrop"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{An overview of oolax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: oolax.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

The *oolax* package

We illustrate the features available from using examples from the [Example analyses from the evd package](evd-vignette.html) vignette.  We choose the evd package [@evd] because it has its own `confint` and `anova` methods, which is useful for comparative purposes.  Similar examples can be performed using the other extreme value packages supported by *oolax*, namely *evir* [@evir], *extRemes* [@extRemes], *fExtremes* [@fExtremes], *ismev* [@ismev] and *texmex* [@texmex].

```{r}
library(oolax)
library(evd)
# An example from the evd::fgev documentation
uvdata <- evd::rgev(100, loc = 0.13, scale = 1.1, shape = 0.2)
M0 <- evd::fgev(uvdata)
M1 <- evd::fgev(uvdata, nsloc = (-49:50)/100)
adj0 <- alogLik(M0)
adj1 <- alogLik(M1)
summary(adj1)
```

Features provided by *chandwich*

## Confidence intervals

```{r}
confint(adj1)
```

## Confidence regions

## Adjusted likelihood ratio tests

```{r}
anova(adj1, adj0)
```


## References

<script type="text/x-mathjax-config">
   MathJax.Hub.Config({  "HTML-CSS": { minScaleAdjust: 125, availableFonts: [] }  });
</script>

