---
title: "Example analyses from the evd package"
author: "Paul Northrop"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Example analyses from the evd package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: oolax.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

@evd

```{r}
library(oolax)
library(evd)
# An example from the evd::fgev documentation
uvdata <- evd::rgev(100, loc = 0.13, scale = 1.1, shape = 0.2)
M0 <- evd::fgev(uvdata)
M1 <- evd::fgev(uvdata, nsloc = (-49:50)/100)
adj0 <- alogLik(M0)
adj1 <- alogLik(M1)
summary(adj1)
```

Objects returned by `alogLik` can be used to produce confidence intervals for individual parameters, confidence regions for pair of parameters and to perform adjusted likelihood ratio tests to compare nested models. See the [An overview of oolax](an-overview-of-oolax-vignette.html) vignette for examples.

## References

<script type="text/x-mathjax-config">
   MathJax.Hub.Config({  "HTML-CSS": { minScaleAdjust: 125, availableFonts: [] }  });
</script>

